\section{The gap in the Level 2 $\rightarrow$ Level 3 escalation}
\label{sec:gap}

ARC-AGI-3 defines the step beyond static reasoning as \emph{agency}: an interactive benchmark scoring a system's efficiency at acquiring new skills through exploration, model formation, goal inference, and planning --- explicitly probing adaptation to ``unknown unknowns'' \citep{arcprize2026,cholletarc2026}. The framework treats this as continuous with the pursuit of universal generality --- the eventual goal being an agent competent across arbitrary, unseen domains.

We propose that the escalation is mis-cut. Between Level~2 and universal agentic AGI there is a distinct, characterizable rung: agents that are \emph{complete within a domain} --- able to reach, autonomously and by closed-loop discovery, any skill or law derivable from the domain's generative rules --- without any claim to cross-domain transfer. We call this \textbf{Domain-Complete Intelligence (\DCI)}, by analogy to Turing-completeness: not ``can do everything,'' but ``can reach everything derivable within this system.''

This is not narrow AI. Narrow AI has \emph{fixed} skill within a domain (Deep Blue plays chess; it does not discover chess-like regularities in data it has never seen). \DCI\ has open-ended \emph{skill acquisition and discovery} within a domain: it proposes hypotheses, simulates, falsifies, and iterates, reaching competences no designer enumerated in advance. By \emph{discovery} we mean, operationally, the capacity to propose, test, falsify, and retain models or skills \emph{not enumerated at design time} --- the property that separates \DCI\ from advanced optimization over a fixed, pre-specified space (a distinction we make precise in \S\ref{sec:cross}). Its generality is complete \emph{relative to the domain's rule system} and empty outside it. Every genuinely superhuman discovery system built to date --- AlphaFold \citep{jumper2021}, AlphaGeometry \citep{trinh2024}, FunSearch \citep{romera2024} --- is domain-complete-with-a-verifier, not general. That is not a limitation of those systems; it is the shape of the rung.
