\section{The verifier is load-bearing}
\label{sec:verifier}

What distinguishes a rule substrate from mere structure is that it closes the discovery loop. In open-world settings there is no oracle, so ``agentic'' behavior degrades into plausible-sounding action with no falsification signal. An executable $\Rule$ converts \emph{agentic} from a vibe into a testable loop: the agent's proposals are adjudicated by the domain itself. This is precisely what universal AGI lacks and cannot cheaply acquire, and precisely what every superhuman discovery system to date has possessed.

The literal demonstration is already in hand. Under a triple-action functional --- trajectory reconstruction ($I_{\max}$), architectural sparsity ($E_{\min}$), and energy-conservation enforcement ($\mathcal{L}_{\mathrm{Symmetry}}$, grounded in Noether's theorem) --- the Minimum-Action Learning (MAL) system selects the correct symbolic force law from a candidate basis on noisy orbital data: it recovers the inverse-square (Kepler) and linear (Hooke) laws, returning the Kepler third-law exponent as $p = 3.01 \pm 0.01$ at roughly a $40\%$ reduction in training energy versus prediction-error-only baselines, with an energy-conservation criterion discriminating the true law where short-term fit alone cannot --- lifting a $40\%/90\%$ direct-selection rate to $100\%$ pipeline-level identification \citep{frasch2026b}. (A wide-stencil preprocessing step cuts noise variance ${\sim}10^{4}\times$, turning an $\mathrm{SNR}\approx 0.02$ problem into a learnable one at ${\approx}\,1.6$.)

Crucially --- and this is a \emph{feature} for the present argument, not a caveat --- the paper is explicit that MAL is a model-\emph{selection} framework over a specified basis, not an open-ended discovery system. This locates MAL precisely: it instantiates the \emph{selection} facet of domain-completeness (reaching the correct law within a given rule library), while the fuller facet --- extending the library \emph{within the domain's own generative grammar} --- is the remaining step. Both are domain-bounded; neither is universal. That boundedness is exactly the \DCI\ signature: completeness \emph{relative to a rule substrate} (here, the basis library and the conservation law that adjudicates it), not open-ended universal discovery. A domain-complete discovery loop closing against a physics substrate is what the Keplerian rung (\S\ref{sec:positioning}) looks like when instantiated --- and its principled boundedness is the point, not a shortfall.
