\section{Related work: verifier-guided search and world models}
\label{sec:related}

\DCI\ names and unifies a pattern already visible in the most successful discovery systems, and locates what they share and what they lack.

\emph{Verifier-guided search.} FunSearch pairs a generator with a systematic evaluator to discover programs that surpass known results in extremal combinatorics and bin-packing \citep{romera2024}; AlphaGeometry couples a neural proposer with a symbolic deduction engine that checks each step \citep{trinh2024}. In both, the evaluator/deduction engine is exactly the executable verifier $\Rule$: an oracle internal to the domain that adjudicates proposals. The refinement loops that defined recent ARC-AGI progress --- per-task iterative program optimization guided by a feedback signal \citep{cholletarc2026} --- are the same closed loop without a domain prior. \DCI's contribution here is not the loop but the claim that adding a \emph{compressed, vertically organizing} prior $\Prin$ is what turns a verifier loop into a domain-\emph{complete} discoverer, and what supplies the object/meta reflexivity (\S\ref{sec:cross}) that a bare evaluator does not.

\emph{World models.} A parallel tradition learns a model of environment dynamics and plans or ``imagines'' within it \citep{ha2018}. \DCI's object level $\Dom$ is a world model in this sense --- but one whose dynamics are constrained by an explicit action principle $\Prin$ rather than fit tabula rasa from data. This is the load-bearing difference: verifiability and closed-loop discovery come from the principle plus an executable substrate, not from data coverage alone, which is why \DCI\ inherits the compression (small $P$-cost, cheap $E$) that a learned world model must pay for in experience. Where program-search systems supply the verifier but no prior, and world-model agents supply a learned model but no verifier-grade ground truth, \DCI\ is the configuration that holds both --- a first-principles model \emph{and} an executable adjudicator --- and is therefore complete within its machine rather than merely competent within it.
